Parsing Postfix Log Files

### 2009-04-17 doc/logparser-other-parsers-review.tex
Corrections from proofreading.
Corrections based on proofreading to the preface and chapters 1-3.

At the start of this project ten Postfix log file parsers were tested, with
the hope of finding a suitable parser to build upon, rather than starting
from scratch.  It was quite difficult to find ten parsers to review for
this project, and the functionality offered by those parsers ranges from
quite basic to much more mature, depending on the needs of the author of
the parser.

It was hoped to reuse an existing parser rather than writing one from
scratch, but the existing parsers considered were rejected for one or more
reasons.  The effort required to adapt and improve an existing parser was
judged to be greater than the effort to write a new one, because the
techniques used by the existing parsers severely limited their potential:
some ignored the majority of log lines, parsing specific log lines
accurately, but without any provision for parsing new or similar log lines;
others sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines of the same category, e.g.\ rejecting a
mail delivery attempt.  The first parser reviewed~\cite{log-mail-analyser}
is the only previously published research in this area that the author is
aware of; their research aims to show that providing the data from log
files in a more accessible form is helpful to systems administrators.

The ten parsers have been reviewed and compared with \parsername{}, this
project's finished parser, to show how much effort would have been required
to use those to fulfil the aims and requirements of this project.  It is
important to compare and contrast newly developed algorithms and parsers
against those already available, to accurately judge what improvements, if
any, are delivered by the newcomers.

Some important differences exist between \parsername{} and most or all of
the parsers reviewed here:

@@

    \item None of the parsers reviewed perform the kind of advanced parsing
        required for this project or deal with the complications described
        in \sectionref{complications}.

    \item Only \parsername{} enables parsing of new log lines without
        extensive and intrusive modifications to the parser; \parsernames{}
        architecture is described in \sectionref{parser architecture}.

    \item The parsers reviewed all produce a report of varying complexity
        and detail, whereas \parsername{} does not; it extracts data and
        leaves generation of reports from the data to other programs.
        Using an \acronym{SQL} database simplifies the process of
        generating such reports (discussed in \sectionref{database as
        API}); some sample queries are given in \sectionref{motivation}.
        The parser developed for this project is designed to enable much
        more detailed log file analysis by providing a stable platform for
        subsequent programs to develop upon.

    \item Most of the reviewed parsers silently ignore log lines they
        cannot handle, whereas \parsername{} complains loudly about every
        single log line it fails to parse.  The exception is AWStats, which
        outputs the percentage of log lines it was unable to parse, but
        does not output the log lines themselves.

    \item A minor difference is that most parsers do not handle compressed
        log files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed log files, but Sawmill exhibits a dramatic
        increase in parsing time when doing so.  Support for reading
        compressed log files is quite helpful, as it dramatically reduces
        the disk space required to store historical log files.

@@

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server \acronym{IP}
address and hostname, HELO hostname, queueid, start time, end time,
\acronym{SMTP} code, enhanced status code, sender, recipient, size, message
ID, delay and delays.

@@

There only appears to be one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \acronym{LMA} is
quite different from \parsername{}: it attempts to present correlated data
from log files in a form suitable for a systems administrator to search
using the myriad of standard Unix text processing utilities already
available.  It produces a \acronym{CSV} file and either a MySQL or Berkeley
DB database.  The decision to support both \acronym{CSV} and Berkeley DB
appears to have been a serious limitation: both formats have limitations
that will be explained later.  Hardly any documentation is provided with
\acronym{LMA}, though some documentation is available
in~\cite{log-mail-analyser}.  Studying the source code is informative,
though this author had difficulty as the authors of \acronym{LMA} wrote in
Italian.

\acronym{CSV} is a very simple format where each record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \acronym{CSV} files include the need to escape separators in
the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  \acronym{CSV} files do not have a
standard mechanism to document the fields or the separator, unlike
\acronym{SQL} databases where every database includes a schema naming the
fields and the type of data they store (e.g.\ integer, text, timestamp).
The \acronym{CSV} record format is not documented, but the output file
contains a comment giving the format:\newline{} \texttt{\# Timestamp|Nome
Client|IP Client|IP Server|From|To|Status|Size} \newline{}\acronym{LMA}
treats \acronym{CSV} lines starting with \texttt{\#} as comments, but not
all \acronym{CSV} parsers will.

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \acronym{SQL} databases that store arbitrary tuples.  In
\acronym{LMA}'s main table the key is an integer referred to by secondary
tables, and the value is a \acronym{CSV} line containing all of the data
for that row.  The secondary by-sender, by-recipient, by-date, and
by-\acronym{IP} tables use the sender/recipient/date/\acronym{IP} address
as the key, and the value is a \acronym{CSV} list of integers referring to
the main table.  This effectively re-implements \acronym{SQL} foreign keys,
but without the functionality offered by even the most basic of
\acronym{SQL} databases (e.g.\ joins, ordering, searches).  It also
requires custom code to search on some combination of the above, though the
authors of \acronym{LMA} did provide some queries: IP-STORY, FROM-STORY,
DAILY-EMAIL, and DAILY-REJECT\@.  Berkeley DB appears to be the least
useful of the three output formats: it does not provide the functionality
of a basic \acronym{SQL} database, and unlike \acronym{CSV} files it cannot
be used with standard Unix text processing tools.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an existing \acronym{SQL} database,
unlike with Berkeley DB\@; all \acronym{SQL} databases embed the schema
into the database and provide commands for displaying it.  Berkeley DB does
not embed a schema in its files, because it does not need one, and
embedding one would not provide any benefit; it only provides \textbf{(key,
value)} pairs, so any additional structuring of the data is imposed by the
application, thus the application must document this structure.  MySQL
support was not tested because the schema required is not documented.

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\acronym{CSV} output, \acronym{LMA} stores the following data: time and
date of the log line, client hostname and \acronym{IP} address, server
\acronym{IP} address, sender and recipient addresses, \acronym{SMTP} code,
and size (for accepted mails only).  Unlike \parsername{} it does not store
the server hostname, HELO hostname, queueid, start and end times,
timestamps for each log line, enhanced status code, delivery delays, or
message id (for accepted mails only).  Handling of multiple recipients,
\acronym{SMTP} codes, or remote servers\footnote{A single mail may be sent
to multiple remote servers if it was addressed to recipients in different
domains, or Postfix needs to try multiple servers for one or more
recipients.} is not explained; experimental observation shows that multiple
records are added when a mail had multiple recipients (sadly the records
are not associated or linked in any way), and presumably the same approach
is taken when there were multiple destination servers.

\acronym{LMA} requires major changes to the parser code to parse new log lines
or to extract additional data.  The code is structured as a long series of
blocks that each handle all log lines matching a single regex, so parsing
new log lines requires modifying an existing regex or carefully inserting a
new block in the correct place; extracting extra data will require
modifying multiple blocks, regexes, or both.

\acronym{LMA} does not deal with any of the complications discussed in
\sectionref{complications}, except for correlating log lines by queueid;
not correlating log lines by pid means it cannot correlate most rejections.
It does not differentiate between different types of rejections, so it is
not suitable for the purposes of this project; the data about which
restriction caused the rejection is discarded, whereas the main goal of
this project is to retain that data to aid optimisation and evaluation of
anti-spam techniques.  \acronym{LMA} fails to parse Postfix log files
generated on Solaris hosts because the fields automatically prepended to
each log line differ from those added on Linux hosts; log files from
Solaris hosts (and possibly other operating systems) thus require
pre-processing before parsing by \acronym{LMA}.  The \numberOFlogFILES{}
pre-processed test log files were parsed without complaint by
\acronym{LMA}, although it produced 32 entries in the output \acronym{CSV}
file for every rejection in the input log file; it also missed some 40\% of
delivered mail.  Once these deficiencies were discovered the author did not
spend any more time checking the results.

\acronym{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  These reports search the Berkeley DB files
for matching records: the first three extract \acronym{CSV} lines for the
specified client \acronym{IP} address, sender address, or date
respectively.  DAILY-REJECT initially failed with an error message from the
Perl interpreter;\footnote{The error messages were:
\newline{}\texttt{Undefined subroutine \&main::LIST called at queryDB.pl
line 372.}\newline{}\texttt{Undefined subroutine \&main::EXTRACT\_FROM\_DB
called at queryDB.pl line 379.}} after correcting the errors in the code it
worked, extracting the \acronym{CSV} lines for the specified day where the
\acronym{SMTP} code signifies a rejection.  All of these reports are
trivially simple to produce from the \acronym{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

The output from DAILY-REJECT and the \texttt{awk} command is not exactly
the same; the author did not spend substantial time attempting to explain
these differences.

@@

    \item The output from DAILY-REJECT is missing some records that are
        present in the \acronym{CSV} file; this may be because it uses the
        Berkeley DB files instead, and there may be differences between the
        contents.

    \item Some records output by DAILY-REJECT are truncated: they are
        missing the last $|$ that separates the fields and the newline
        following it, so the line containing only a single space is
        concatenated with the record.

@@

In summary, \acronym{LMA} appears to be a proof of concept, written to
demonstrate the point of their paper (that having this information in an
accessible fashion is useful to systems administrators), rather than a
program designed to be useful in a production environment.

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% DNS name and \acronym{IP} address host, mail server \acronym{IP} address,
% sender, receiver and e-mail status (sent, rejected).  Undocumented
% schema.  Design decision to use \acronym{CSV} as an intermediate format
% between the log file and the database seems to have been restrictive.
% Appears to require a queueid but majority of log lines (e.g.\ rejections)
% lack a queueid.  Supports whitelisting \acronym{IP} addresses when
% parsing logs, but whitelisting when generating reports/data mining would
% be preferable.  Supporting Berkeley DB is probably limiting the software;
% an example is the difficulty in searching a pipe-delimited string, so
% they have re-implemented foreign keys with tables keyed by ip address and
% so on pointing at the main table - this also will not scale well.  There
% does not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any data, and it does not
extract any data that it does not require to produce its report; e.g.\ it
does not extract the HELO hostname, queueid, start and end times,
timestamps for each log line, or message id.  Both the parsing and
reporting are difficult to extend because it is a specialised tool, unlike
the easily extensible design of \parsername{}.  It does not correlate log
lines by queueid or \acronym{pid}, and does not need to deal with the
complications encountered during this project (\sectionref{complications}).
Pflogsumm produces a useful report, and successfully parsed the
\numberOFlogFILES{} log files it was tested with.\footnote{The results it
reported were not verified in detail, but it did not report any errors, and
has an excellent reputation amongst Postfix users.}  Pflogsumm has many
options to include or exclude certain sections of the report, all clearly
documented; by default it includes the following:

@@

    \item For received mail: per-domain totals for mails sent, deferred,
        average delay, maximum delay, and bytes delivered.  For received
        mail: per-domain totals for size and number of mails received.

@@

    \item Summary of why mail delivery was deferred or failed, why mails
        were bounced, why mails were rejected, and warning messages.

@@

Pflogsumm is the only reviewed parser that distinguishes between different
rejections.

@@

Sawmill is a general purpose commercial product that parses 818 log file
formats (as of 2008/11/23) and produces reports from the extracted data.
Its data extraction facilities (described later) are too limited to save
enough data for the purposes of this project: although it can extract three
different sets of data from Postfix log files, they are not interlinked in
any way.  The documentation does not suggest that Sawmill correlates log
lines by either queueid or pid or deals with the other difficulties
documented in \sectionref{complications}.

@@

    \item \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/postfix.html}{2008/11/23}.
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \acronym{IP} address, and
        virus.  It also counts the number of and total size of all mails
        delivered.  The fields \texttt{server}, \texttt{uid},
        \texttt{relay}, and \texttt{virus} are not explained in the
        documentation: \texttt{server} is probably the hostname or
        \acronym{IP} address of the server the mail is delivered to;
        \texttt{relay} might be the delivery method: \acronym{SMTP}, local
        delivery, or \acronym{LMTP}; \texttt{uid} might be the uid of the
        user submitting mail locally.  Postfix does not do any form of
        virus checking (though it has many options for cooperating with an
        external virus scanner), so the \texttt{virus} field is a mystery.

    \item \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/postfix_ii.html}{2008/11/23}.
        Fields extracted: from, to, RBL list, client hostname, and client
        \acronym{IP} address.  It also counts the number and total size of
        all mails delivered.

    \item \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/beta_postfix.html}{2008/11/23}.
        Fields extracted: from, to, client hostname, client \acronym{IP}
        address, relay hostname, relay \acronym{IP} address, status,
        response code, RBL list, and message id.  It also counts the number
        and size of all mails delivered, processed, blocked, expired, and
        bounced.

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, and all searches produced accurate results.  The interface for
searching is neither as simple to use nor as informative as the interface
provided by Splunk (see \sectionref{Splunk review}).  The administrative
interface is much easier to use than Splunk's: it took only five minutes to
start parsing a whole directory of log files.

When tested with the \numberOFlogFILES{} test log files it performed
adequately, though the rate it processed log files at did slow down
noticeably as it progressed.  Sawmill supports reading compressed log files
but it exhibits a dramatic slow down when doing so: it took six hours to
parse the first half of the log files, and twelve hours to parse the next
third; after twenty four hours parsing the remaining sixth it crashed due
to lack of disk space.  On the second parsing attempt the log files were
uncompressed beforehand and parsing took eight hours.

In summary, Sawmill suffers from supporting so many types of log files: it
parses many types of log files, but none of them very well; it is probably
much more useful when parsing log files where each log line is
self-contained (e.g.\ web server log files), rather than log files
containing interlinked log lines.  It is not suitable as a base for this
parser, as the source code made available is encrypted and not intended
for modification; in addition the architecture would probably need to be
overhauled or replaced to deal with correlating log lines.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields extracted by Splunk
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The data store is not available for use by external programs,
whereas \parsername{} provides the database and leaves it to the user to
utilise it without limit or restriction.  The web interface is optimised
for interactive use rather than automated queries, and it does not appear
to be possible to write independent tools to utilise the Splunk database.
Some additional Postfix reports are supposedly available on
\urlLastChecked{http://www.splunkbase.com/}{2008/11/23}, but the author was
unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though most are variations of a bar or pie chart, except the bubble and
heatmap graphs.  It is trivial to drill down through the graphs to select a
portion of the data (e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address).  All searches performed using the indexed data returned
reasonable results.

>>>>

At the start of this project ten Postfix log file parsers were tested, with
the hope of finding a suitable parser to build upon, rather than writing
one from scratch.  It was quite difficult to find ten parsers to review for
this project, and the functionality offered by those parsers ranges from
quite basic to much more mature, depending on the needs of the author of
the parser.

It was hoped to reuse an existing parser rather than writing one from
scratch, but the existing parsers considered were rejected for one or more
reasons.  The effort required to adapt and improve an existing parser was
judged to be greater than the effort to write a new one, because the
techniques used by the existing parsers severely limited their potential:
some ignored the majority of log lines, parsing specific log lines
accurately, but without any provision for parsing new or similar log lines;
others sloppily parsed the majority of log lines, but were incapable of
distinguishing between log lines within one category, e.g.\ not
distinguishing between different anti-spam techniques causing delivery
attempts to be rejected.  The first parser
reviewed~\cite{log-mail-analyser} is the only previously published research
in this area that the author is aware of; that research aims to show that
providing the data from log files in a more accessible form is helpful to
systems administrators.

The ten parsers have been reviewed and compared with \parsername{}, this
project's parser, to show how much effort would have been required to use
one of those parsers to fulfil the aims and requirements of this project.
It is important to compare and contrast newly developed algorithms and
parsers against those already available, to accurately judge what
improvements, if any, are delivered by the newcomers.

Some important differences exist between \parsername{} and most or all of
the parsers reviewed in this chapter:

@@

    \item None of the reviewed parsers perform the kind of advanced parsing
        required for this project, or deal with the complications described
        in \sectionref{complications}.

    \item Only \parsername{} enables parsing of new log lines without
        extensive and intrusive modifications to the parser; the
        architecture that enables this is documented in \sectionref{parser
        architecture}.

    \item The reviewed parsers all produce a report of varying complexity
        and detail, whereas \parsername{} does not; it extracts data and
        leaves generation of reports to other programs.  Using a
        \acronym{SQL} database simplifies the process of generating such
        reports (discussed in \sectionref{database as API}); some sample
        queries are given in \sectionref{motivation}.  The parser developed
        for this project is designed to enable much more detailed log file
        analysis by providing a stable platform for subsequent programs to
        build upon.

    \item Most of the reviewed parsers silently ignore log lines they
        cannot handle, whereas \parsername{} warns about every single log
        line it fails to recognise.  The exception is AWStats, which
        outputs the percentage of log lines it was unable to parse, but
        does not output the log lines themselves.

    \item A minor difference is that most parsers do not handle compressed
        log files; both \parsername{} and Splunk handle them transparently,
        without user intervention; Sawmill and Lire can be configured to
        support compressed log files, but Sawmill exhibits a dramatic
        increase in parsing time when doing so.  Support for reading
        compressed log files is quite helpful, because it dramatically
        reduces the disk space required to store historical log files.

@@

The data extracted by \parsername{} is documented in
\sectionref{connections table} and \sectionref{results table}; for
convenience that list is repeated here: client and server IP address and
hostname, HELO hostname, queueid, start time, end time, \acronym{SMTP}
code, enhanced status code, sender, recipient, size, message-id, delay, and
delays.

@@

There appears to be only one prior published paper about parsing Postfix
log files: \textit{Log Mail Analyzer: Architecture and Practical
Utilizations\/}~\cite{log-mail-analyser}.  The aim of \acronym{LMA} is
quite different from \parsername{}: it attempts to present correlated data
from log files in a form suitable for a systems administrator to search
using the myriad of standard Unix text processing utilities already
available.  It produces a \acronym{CSV} file and either a MySQL or Berkeley
DB database.  Hardly any documentation is provided with \acronym{LMA}, but
some documentation is available in~\cite{log-mail-analyser}.  Studying the
source code is informative, though this author had difficulty as the
authors of \acronym{LMA} wrote in Italian.

\acronym{CSV} is a very simple format where each record is stored in a
single line, with fields separated by a comma or other punctuation symbol.
Problems with \acronym{CSV} files include the need to escape separators in
the data stored, providing multiple values for a field (e.g.\ multiple
recipients), and adding new fields.  \acronym{CSV} files do not have a
standard mechanism to document the fields or the separator, unlike
\acronym{SQL} where every database includes a schema naming the fields and
the type of data they store (e.g.\ integer, text, timestamp).  The
\acronym{CSV} record format used by \acronym{LMA} is not documented, but
the output file contains a comment giving the format:\newline{} \texttt{\#
Timestamp|Nome Client|IP Client|IP Server|From|To|Status|Size}
\newline{}\acronym{LMA} treats \acronym{CSV} lines starting with
\texttt{\#} as comments, but not all \acronym{CSV} parsers will.

Berkeley DB only supports storing simple \textbf{(key, value)} pairs,
unlike \acronym{SQL} databases that store arbitrary tuples.  In
\acronym{LMA}'s main table the key is an integer referred to by secondary
tables, and the value is a \acronym{CSV} line containing all of the data
for that row.  The secondary by-sender, by-recipient, by-date, and by-IP
tables use the sender/recipient/date/IP address as the key, and the value
is a \acronym{CSV} list of integers referring to the keys in the main
table.  This effectively re-implements \acronym{SQL} foreign keys, but
without the functionality offered by even the most basic of \acronym{SQL}
databases, e.g.\ joins, ordering, searches.  It also requires custom code
to search on a combination of attributes, though the authors of
\acronym{LMA} did provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL, and DAILY-REJECT (all described later).  Berkeley DB appears
to be the least useful of the three output formats: it does not provide the
functionality of a basic \acronym{SQL} database, and unlike \acronym{CSV}
files it cannot be used with standard Unix text processing tools.

The schema used with the MySQL database is undocumented, but at least it is
possible to discover the schema with an existing \acronym{SQL} database,
unlike with Berkeley DB\@; all \acronym{SQL} databases embed the schema
into the database and provide commands for displaying it.  Berkeley DB does
not embed a schema in its files, because it does not need one, and
embedding one would not provide any benefit; it only provides \textbf{(key,
value)} pairs, so any additional structuring of the data is imposed by the
application, and it is the responsibility of the author of the application
to document this structure.  MySQL support was not tested because the
schema is not documented, so the required tables could not be created.

Whether a MySQL database or Berkeley DB table is chosen in addition to the
\acronym{CSV} output, \acronym{LMA} stores the following data: time and
date of the log line, client hostname and IP address, server IP address,
sender and recipient addresses, \acronym{SMTP} code, and size (for accepted
mails only).  Unlike \parsername{} it does not store the server hostname,
HELO hostname, queueid, start and end times, timestamps for each log line,
enhanced status code, delivery delays, or message id (for accepted mails
only).  Handling of multiple recipients, \acronym{SMTP} codes, or remote
servers\footnote{A single mail may be sent to multiple remote servers if it
was addressed to recipients in different domains, or Postfix needs to try
multiple servers for one or more recipients.} is not explained;
experimental observation shows that multiple records are added when a mail
had multiple recipients, but the records are not associated or linked in
any way, and presumably the same approach was taken when there were
multiple destination servers.

\acronym{LMA} requires major changes to the parser code to parse new log lines
or to extract additional data.  The code is structured as a long series of
blocks that each handle all log lines matching a single regex, so parsing
new log lines requires modifying an existing regex or carefully inserting a
new block in the correct place; extracting extra data would require
modifying multiple blocks, regexes, or both.

\acronym{LMA} does not deal with any of the complications discussed in
\sectionref{complications}, except for correlating log lines by queueid;
not correlating log lines by \acronym{pid} means it cannot correlate most
rejected delivery attempts.  It does not differentiate between different
types of rejections, so it is not suitable for the purposes of this
project; the data about which restriction caused the rejection is
discarded, whereas the main goal of this project is to retain that
information to aid optimisation and evaluation of anti-spam techniques.
\acronym{LMA} fails to parse Postfix log files generated on Solaris hosts
because the fields automatically prepended to each log line differ from
those prepended on Linux hosts; log files from Solaris hosts (and possibly
other operating systems) thus require pre-processing before parsing by
\acronym{LMA}.  The \numberOFlogFILES{} pre-processed test log files were
parsed without complaint by \acronym{LMA}, although it produced 32 entries
in the output \acronym{CSV} file for every rejection in the input log
files; it also missed some 40\% of delivered mail.  Once these deficiencies
were discovered the author did not spend any more time checking the
results.

\acronym{LMA} does provide some simple reports: IP-STORY, FROM-STORY,
DAILY-EMAIL and DAILY-REJECT\@.  These reports search the Berkeley DB files
for matching records: the first three produce \acronym{CSV} lines for the
specified client IP address, sender address, or date respectively.
DAILY-REJECT initially failed with an error message from the Perl
interpreter;\footnote{The error messages were: \newline{}\texttt{Undefined
subroutine \&main::LIST called at queryDB.pl line
372.}\newline{}\texttt{Undefined subroutine \&main::EXTRACT\_FROM\_DB
called at queryDB.pl line 379.}} after correcting the errors in the code it
worked, producing \acronym{CSV} lines for the specified day where the
\acronym{SMTP} code signifies a rejection.  All of these reports are
trivially simple to produce from the \acronym{CSV} file using the standard
Unix tool \texttt{awk}\glsadd{awk}; the most complicated, DAILY-REJECT, is
merely:

@@

The output from DAILY-REJECT and the \texttt{awk} command is not exactly
the same; the author did not spend substantial time attempting to rectify
these differences.

@@

    \item The output from DAILY-REJECT is missing some records that are
        present in the \acronym{CSV} file; this may indicate differences
        between the data stored in the \acronym{CSV} and Berkeley DB files.

    \item Some records output by DAILY-REJECT are truncated: they are
        missing the last $|$ that separates fields and also the newline
        following it, so the line containing only a single space is
        concatenated with the record.

@@

In summary, \acronym{LMA} appears to be a proof of concept, written to
demonstrate the point of their paper, that making the information contained
in log files available in an accessible fashion is useful to systems
administrators, rather than a program intended to be useful in a production
environment.

% Literature review notes:
%
% Hard-coded parsing, requiring code changes to add more.  Attempts to
% correlate log lines, saves data to database for data mining purposes.
% Hard to extend/expand/understand.  Appears to only save: date and hour,
% DNS name and IP address host, mail server IP address, sender, receiver
% and e-mail status (sent, rejected).  Undocumented schema.  Design
% decision to use \acronym{CSV} as an intermediate format between the log
% file and the database seems to have been restrictive.  Appears to require
% a queueid but majority of log lines (e.g.\ rejections) lack a queueid.
% Supports whitelisting IP addresses when parsing log filess, but
% whitelisting when generating reports/data mining would be preferable.
% Supporting Berkeley DB is probably limiting the software; an example is
% the difficulty in searching a pipe-delimited string, so they have
% re-implemented foreign keys with tables keyed by ip address and so on
% pointing at the main table - this also will not scale well.  There does
% not appear to be any attempt to deal with the complications I have
% encountered: their parsing is not detailed enough to encounter them.  It
% does not run properly; does not create any output; throws up errors.

@@

Pflogsumm produces a report designed for troubleshooting rather than
in-depth analysis.  It does not support saving any of the data it extracts
from log files, and it does not extract any data that it does not require
to produce its report: the HELO hostname, queueid, start and end times,
timestamps for each log line, or message id.  Both the parsing and
reporting are difficult to extend because it is a specialised tool, unlike
the easily extensible design of \parsername{}.  It does not correlate log
lines by queueid or \acronym{pid}, and does not need to deal with the
complications encountered during this project (\sectionref{complications}).
Pflogsumm produces a useful report, and successfully parsed the
\numberOFlogFILES{} log files it was tested with.  The results it reported
were not verified in detail, but it did not report any errors, and has an
excellent reputation amongst Postfix users.  Pflogsumm has many options to
include or exclude certain sections of the report, all clearly documented;
by default its report includes the following:

@@

    \item For received mail: per-domain totals for the number of mails
        sent, number of mails deferred, average delay, maximum delay, and
        bytes delivered.  For received mail: per-domain totals for size and
        number of mails received.

@@

    \item Summary of why mail delivery was deferred or failed, why mails
        were bounced, why mails were rejected, and warning messages from
        Postfix.

@@

Pflogsumm is the only reviewed parser that distinguishes between different
rejected delivery attempts, an important requirement for this project.

@@

Sawmill is a general purpose commercial product that parses 818 log file
formats (as of 2008/11/23) and produces reports from the extracted data.
Its data extraction facilities (described later) are too limited to save
enough data for the purposes of this project: although it can extract three
different sets of data from Postfix log files, they are not linked in any
way.  The documentation does not suggest that Sawmill correlates log lines
by either queueid or \acronym{pid}, or deals with the other difficulties
documented in \sectionref{complications}.

@@

    \item
        \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/postfix.html}{2008/11/23}.
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin IP address, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the hostname or IP address of the
        server the mail is delivered to; \texttt{relay} might be the
        delivery method: \acronym{SMTP}, local delivery, or \acronym{LMTP};
        \texttt{uid} might be the uid of the user submitting mail locally.
        Postfix does not do any form of virus checking (though it has many
        options for cooperating with an external virus scanner), so the
        \texttt{virus} field is a mystery.

    \item
        \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/postfix_ii.html}{2008/11/23}.
        Fields extracted: from, to, RBL list, client hostname, and client
        IP address.  It also counts the number and total size of all mails
        delivered.

    \item
        \urlLastChecked[\hfill{}\newline{}]{http://www.thesawmill.co.uk/formats/beta_postfix.html}{2008/11/23}.
        Fields extracted: from, to, client hostname, client IP address,
        relay hostname, relay IP address, status, response code, RBL list,
        and message id.  It also counts the number and size of all mails
        delivered, processed, blocked, expired, and bounced.

@@

Sawmill's web interface supports searching on any combination of the fields
it extracts, and all searches produced accurate results.  The interface for
searching is neither as simple to use nor as informative as the interface
provided by Splunk (see \sectionref{Splunk review}).  However, the
administrative interface is much easier to use than Splunk's: it took only
five minutes to start parsing all of the log files in a directory.

When tested with the \numberOFlogFILES{} test log files it performed
adequately, though the rate it processed log files at did slow down
noticeably as it progressed.  Sawmill supports reading compressed log files
but it exhibits a dramatic slow down when doing so: it took six hours to
parse the first half of the log files, and twelve hours to parse the next
third; after twenty four hours spent parsing the remaining sixth it crashed
due to lack of disk space.  On the second parsing attempt the log files
were uncompressed beforehand and parsing took eight hours.

In summary, Sawmill suffers from supporting so many types of log files: it
is probably much more useful when parsing log files where each log line is
self-contained (e.g.\ web server log files), rather than log files where
data is spread across multiple log lines.  It is not suitable as a base for
this parser, because the source code made available is encrypted and not
intended for modification; in addition the architecture would probably need
to be overhauled or replaced to deal with correlating log lines.

@@

Splunk aims to index all of an organisation's log files, providing a
centralised view capable of searching and correlating diverse log sources.
The web interface provides search functionality, generating statistics and
graphs in real time, a facility not provided by \parsername{}.  Splunk
allows quite complicated searches, based on the fields it extracts
(described later) or the full text of the log line, though it is not
possible to search on partial words.  Search results are not available for
use with other tools.  Searches can be saved for reuse; saved searches can
be run periodically and the results mailed to a recipient or sent to an
external program for further processing.  The author was unable to save
searches, though that may have been because of limitations in the free
version.  The web interface is optimised for interactive use rather than
automated queries, and it does not appear to be possible to write
independent tools to utilise the Splunk database, whereas \parsername{}
provides the database and leaves it to the user to utilise it without limit
or restriction.  Some additional Postfix reports are supposedly available
on \urlLastChecked{http://www.splunkbase.com/}{2008/11/23}, but the author
was unable to find any Postfix reports, or indeed reports for any other log
file types: every category was empty, even those that the web site claimed
had numerous reports available.  Many types of graphs can be generated,
though all except the bubble and heatmap graphs are variations of a bar or
pie chart.  It is trivial to drill down through the graphs to select a
portion of the data, e.g.\ select the hour with the largest number of
events, then select a particular host, and finally a specific sender
address.  All searches performed using the indexed data returned reasonable
results.  The full power of \acronym{SQL} is available when searching the
data extracted by \parsername{}, allowing the user to search on arbitrarily
complicated conditions.

The web interface is quite attractive and simple to use when searching, but
as an administrator it seems unnecessarily difficult to perform simple
tasks.  When testing Splunk it took roughly 30 minutes to figure out how to
add a single log file to be indexed for later searching, with the added
downside that the log file was copied into a spool directory before
indexing, doubling the disk space usage.  The next test was to index all
the log files in a particular directory, but after three hours, reading all
the available documentation, and numerous futile attempts, the author was
still unable to index all the log files in a directory using the web
interface.  Using Splunk's \acronym{CLI} rather than the web interface was
more successful: the command ``\texttt{splunk find logs
}\textit{log-directory\/}'' added 40 of the \numberOFlogFILES{} log files
to the queue for indexing.  Further attempts enqueued the same 40 log
files, without explaining why the others were excluded.\footnote{The log
files that were added multiple times appear to have been indexed once only;
presumably Splunk keeps track of the log files it has indexed and discards
requests to index log files for a second time.  This may or may not be a
useful feature for \parsername{}.}  The command did not have an option to
ensure the log files would be processed in the order they were created,
though such an option may be neither necessary nor beneficial with Splunk.
Subsequently the author was successful in adding a single log file at a
time using the command ``\texttt{splunk add tail }\textit{filename\/}''.  A
simple loop using that command was enough to add all the desired log files.
Splunk will periodically check all indexed log files for updates unless
they are manually removed from its list; this may or may not be useful
behaviour.  Splunk did not appear to have any difficulty in indexing the
log files once they had been successfully added to its queue.
\parsername{} parses the log files it is instructed to parse, in the order
they are given; periodic parsing of log files is a task an administrator
can easily achieve with \texttt{cron(8)} and \texttt{logrotate(8)}.

Copious documentation is made available on
\urlLastChecked{http://www.splunk.com/}{2008/11/23}, but the abundance of
material and lack of organisation makes it hard to find the topic being
sought, and searches confusingly tended to return results from old
documentation rather than new.  In general the documentation appears to
have been written by someone intimately acquainted with the software, who
has difficulty understanding how a newcomer would approach tasks or the
questions they would ask.

Splunk supports reading compressed log files without any extra
configuration by the user, like \parsername{}.  The free version of Splunk
limits the volume of data indexed per day to 500MB, though a trial
Enterprise licence is available that allows indexing of up to 5GB of data
per day.  In 2007, the cheapest licenced version cost \$5000 plus \$1000
support per annum, and limited the volume of data indexed per day to
500MB\@.  Prices were removed from the Splunk website during 2008; now
Splunk's sales team must be asked for a quote.  Median log file size for
the log files used when evaluating \parsername{} is
\input{build/include-median-file-size}.
